Analysis 2

### 2022-12-04 Analysis 2.qmd
Updates

```{r}
library(tidyverse)
library(broom)
library(terra)
```

Read training and test data and perform any cleaning steps

@@

# Remove ID column only from the training set
housesTrain = select(housesTrain, -Id)

@@

# Remove any of the columns that had missing values
housesTrain =  housesTrain |>
  select(-starts_with(c(colsToDrop, colsToDrop2)))
housesTest =  housesTest |>
  select(-starts_with(c(colsToDrop, colsToDrop2)))

hist(housesTrain$LotArea)

# From the plot below, LotArea should likely cut any value off at 30,000 sq ft
housesTrain$LotArea = clamp(housesTrain$LotArea, lower = 0, upper = 30000)
housesTest$LotArea = clamp(housesTest$LotArea, lower = 0, upper = 30000)

# From the plot below, GrLivArea should likely cut any value off at 4000 sq ft
housesTrain$GrLivArea = clamp(housesTrain$GrLivArea, lower = 0, upper = 4000)
housesTest$GrLivArea = clamp(housesTest$GrLivArea, lower = 0, upper = 4000)

# From the plot below, MSSubClass looks to have a quadratic relationship

# From the plot below, LotArea should should have a log relationship, which means each value should be greater than or equal to 1
housesTrain$LotArea = clamp(housesTrain$LotArea, lower = 1)
housesTest$LotArea = clamp(housesTest$LotArea, lower = 1)

# From the plot below, OpenPorchSF should should have a log relationship, which means each value should be greater than or equal to 1
#housesTrain$OpenPorchSF = clamp(housesTrain$OpenPorchSF, lower = 1)
#housesTest$OpenPorchSF = clamp(housesTest$OpenPorchSF, lower = 1)

# From the plot below, MiscVal should should have a log relationship, which means each value should be greater than or equal to 1
housesTrain$MiscVal = clamp(housesTrain$MiscVal, lower = 1)
housesTest$MiscVal = clamp(housesTest$MiscVal, lower = 1)

# From the plot below, PoolArea should should have a log relationship, which means each value should be greater than or equal to 1
#housesTrain$PoolArea = clamp(housesTrain$PoolArea, lower = 1)
#housesTest$PoolArea = clamp(housesTest$PoolArea, lower = 1)

# Screen for outliers by looking at the histograms of any variables that are all numeric
housesTrain |> 
  select(where(is.numeric)) |>
  gather() |>
  ggplot(aes(x = value)) + 
    facet_wrap(~key, scales = "free") + 
    geom_histogram()

@@

# The sale price has a wide variance. Applying a log transformation helps
housesTrain |>
  ggplot(aes(x = log(SalePrice))) +
  geom_histogram()

# Log of SalePrice should be used to reduce the variance
housesTrain$SalePrice = log(housesTrain$SalePrice)

@@

Full model creation

```{r}
housesTrain |> count(`3SsnPorch`)

fit = lm(SalePrice ~ I(MSSubClass^2) + log(LotArea) + Street + LotShape + LandContour + LotConfig + LandSlope + Neighborhood + Condition1 + Condition2 + BldgType + HouseStyle + OverallQual + OverallCond + YearBuilt + YearRemodAdd + RoofStyle + RoofMatl + ExterQual + ExterCond + Foundation + Heating + HeatingQC + CentralAir + `1stFlrSF` + `2ndFlrSF` + LowQualFinSF + GrLivArea + FullBath + HalfBath + BedroomAbvGr + KitchenAbvGr + TotRmsAbvGrd + Fireplaces + PavedDrive + WoodDeckSF + OpenPorchSF + EnclosedPorch + `3SsnPorch` + ScreenPorch + PoolArea + log(MiscVal) + MoSold + YrSold + SaleCondition, data = housesTrain)
summary(fit)

@@

Test the basic model against the test data set and submit to Kaggle to see the score

```{r}
results = housesTest
results$SalePrice = exp(predict.lm(fit, newdata = housesTest))
results = results |>
  select(Id, SalePrice)

# Export
write_csv(results, "test_submission_10.csv")
```

Try using forward stepping approach

```{r}
# Define intercept-only model, this is also just the median sale price
intercept_only = lm(SalePrice ~ 1, data = housesTrain)

# Define model with all predictors
all = lm(SalePrice ~ ., data = housesTrain)

# Perform forward step-wise regression
forward = step(intercept_only, direction = 'forward', scope = formula(all), trace = 0)

# View final model
summary(forward)

@@

Test the forward model against the test data set and submit to Kaggle

```{r}
results = housesTest
results$SalePrice = exp(predict.lm(forward, newdata = housesTest))
results = results |>
  select(Id, SalePrice)

# Export
write_csv(results, "test_submission_forward_1.csv")
```

Try using backward stepping approach

```{r}
# Define model with all predictors
all = lm(SalePrice ~ ., data = housesTrain)

# Perform forward step-wise regression
backward = step(all, direction = 'backward', scope = formula(all), trace = 0)

# View final model
summary(backward)

@@

Test the backward model against the test data set and submit to Kaggle

```{r}
results = housesTest
results$SalePrice = exp(predict.lm(backward, newdata = housesTest))
results = results |>
  select(Id, SalePrice)

# Export
write_csv(results, "test_submission_backward_1.csv")
```

Try using stepwise stepping approach

```{r}
# Define intercept-only model, this is also just the average fat amount
intercept_only = lm(SalePrice ~ 1, data = housesTrain)

# Define model with all predictors
all = lm(SalePrice ~ ., data = housesTrain)

# Perform forward step-wise regression
both = step(intercept_only, direction = 'both', scope = formula(all), trace = 0)

# View final model
summary(both)
```

Test the stepwise model against the test data set and submit to Kaggle

```{r}
results = housesTest
results$SalePrice = exp(predict.lm(both, newdata = housesTest))
results = results |>
  select(Id, SalePrice)

# Export
write_csv(results, "test_submission_stepwise_1.csv")
```

>>>>

```{r}
#library(MASS)
library(olsrr)
library(tidyverse)
library(terra)

library(broom)
library(caret)
```

Read training and test data, then drop any columns containing missing values

@@

# Remove ID column only from the training set
housesTrain = dplyr::select(housesTrain, -Id)

@@

# Remove any of the columns that had missing values
housesTrain =  housesTrain |>
  dplyr::select(-starts_with(c(colsToDrop, colsToDrop2)))
housesTest =  housesTest |>
  dplyr::select(-starts_with(c(colsToDrop, colsToDrop2)))
```

```{r}
# Screen for numerical variables that should be log-transformed
housesTrain |> 
  dplyr::select(where(is.numeric)) |>
  gather() |>
  ggplot(aes(x = value)) + 
    facet_wrap(~key, scales = "free") + 
    geom_histogram()
```

The following variables should be log transformed:
- LotArea
- FirstFlrSF
- SecondFlrSF
- GrLivArea
- WoodDeckSF
- OpenPorchSF
- EnclosedPorch
- ThirdSsnPorch
- ScreenPorch
- PoolArea
- MiscVal
- SalePrice

```{r}

@@

# First ensure that any variables that need log transformations are set to a minimum value of 1 instead of 0. Then, take the log of the values.
housesTrain = housesTrain |>
  mutate(
    LotArea = log(clamp(LotArea, lower = 1)),
    FirstFlrSF = log(clamp(`1stFlrSF`, lower = 1)),
    SecondFlrSF = log(clamp(`2ndFlrSF`, lower = 1)),
    GrLivArea = log(clamp(GrLivArea, lower = 1)),
    WoodDeckSF = log(clamp(WoodDeckSF, lower = 1)),
    OpenPorchSF = log(clamp(OpenPorchSF, lower = 1)),
    EnclosedPorch = log(clamp(EnclosedPorch, lower = 1)),
    ThirdSsnPorch = log(clamp(`3SsnPorch`, lower = 1)),
    ScreenPorch = log(clamp(ScreenPorch, lower = 1)),
    PoolArea = log(clamp(PoolArea, lower = 1)),
    MiscVal = log(clamp(MiscVal, lower = 1)),
    SalePrice = log(clamp(SalePrice, lower = 1)),
  )

housesTest = housesTest |>
  mutate(
    LotArea = log(clamp(LotArea, lower = 1)),
    FirstFlrSF = log(clamp(`1stFlrSF`, lower = 1)),
    SecondFlrSF = log(clamp(`2ndFlrSF`, lower = 1)),
    GrLivArea = log(clamp(GrLivArea, lower = 1)),
    WoodDeckSF = log(clamp(WoodDeckSF, lower = 1)),
    OpenPorchSF = log(clamp(OpenPorchSF, lower = 1)),
    EnclosedPorch = log(clamp(EnclosedPorch, lower = 1)),
    ThirdSsnPorch = log(clamp(`3SsnPorch`, lower = 1)),
    ScreenPorch = log(clamp(ScreenPorch, lower = 1)),
    PoolArea = log(clamp(PoolArea, lower = 1)),
    MiscVal = log(clamp(MiscVal, lower = 1))
  )
```

Forward step regression
```{r}
forwardResults = ols_step_forward_aic(lm(SalePrice ~ ., data = housesTrain))
forwardModel = forwardResults$model

plot(forwardResults)

summary(forwardModel)

results = housesTest
results$SalePrice = exp(predict.lm(forwardModel, newdata = housesTest))
results = results |>
  dplyr::select(Id, SalePrice)

# Export
write_csv(results, "Submissions/Forward_Model.csv")

@@

Backward step regression
```{r}
backwardResults = ols_step_backward_aic(lm(SalePrice ~ ., data = housesTrain))
backwardModel = backwardResults$model

plot(backwardResults)

summary(backwardModel)

results = housesTest
results$SalePrice = exp(predict.lm(backwardModel, newdata = housesTest))
results = results |>
  dplyr::select(Id, SalePrice)

# Export
write_csv(results, "Submissions/Backward_Model.csv")

@@

Stepwise regression
```{r}
stepResults = ols_step_both_aic(lm(SalePrice ~ ., data = housesTrain), details = TRUE)
stepModel = lm(SalePrice ~ OverallQual + GrLivArea + Neighborhood + OverallCond + YearBuilt + LotArea + RoofMatl + KitchenAbvGr + SaleCondition + Condition2 + Foundation + Fireplaces + Heating + ExterQual + Condition1 + PoolArea + ScreenPorch + WoodDeckSF + HeatingQC + CentralAir + BedroomAbvGr + FirstFlrSF + `2ndFlrSF` + Street + LandSlope + HalfBath + EnclosedPorch + SecondFlrSF + MiscVal + PavedDrive + BldgType + `1stFlrSF` + ExterCond + LotConfig + YearRemodAdd, data = housesTrain)

plot(stepResults)

summary(stepModel)

results = housesTest
results$SalePrice = exp(predict.lm(stepModel, newdata = housesTest))
results = results |>
  select(Id, SalePrice)

# Export
write_csv(results, "Submissions/Both_Model.csv")

@@

Best custom model using caret
```{r}
customModel = lm(SalePrice ~ I(MSSubClass^2) + log(LotArea) + Street + LotShape + LandContour + LotConfig + LandSlope + Neighborhood + Condition1 + Condition2 + BldgType + HouseStyle + OverallQual + OverallCond + YearBuilt + YearRemodAdd + RoofStyle + RoofMatl + ExterQual + ExterCond + Foundation + Heating + HeatingQC + CentralAir + `1stFlrSF` + `2ndFlrSF` + LowQualFinSF + GrLivArea + FullBath + HalfBath + BedroomAbvGr + KitchenAbvGr + TotRmsAbvGrd + Fireplaces + PavedDrive + WoodDeckSF + OpenPorchSF + EnclosedPorch + `3SsnPorch` + ScreenPorch + PoolArea + log(MiscVal) + MoSold + YrSold + SaleCondition, data = housesTrain)
summary(fit)

#control = trainControl(method = "LOOCV", number = 10)
control = trainControl(method = "cv")
customModel = train(SalePrice ~ ., data = housesTrain, method = "lm", trControl = control)
summary(customModel)

results = housesTest
results$SalePrice = exp(predict(customModel, newdata = housesTest))
results = results |>
  select(Id, SalePrice)

# Export
write_csv(results, "Submissions/Custom_Model.csv")

@@

Best custom model using external cross validation
```{r}
customModel = lm(SalePrice ~ I(MSSubClass^2) + log(LotArea) + Street + LotShape + LandContour + LotConfig + LandSlope + Neighborhood + Condition1 + Condition2 + BldgType + HouseStyle + OverallQual + OverallCond + YearBuilt + YearRemodAdd + RoofStyle + RoofMatl + ExterQual + ExterCond + Foundation + Heating + HeatingQC + CentralAir + `1stFlrSF` + `2ndFlrSF` + LowQualFinSF + GrLivArea + FullBath + HalfBath + BedroomAbvGr + KitchenAbvGr + TotRmsAbvGrd + Fireplaces + PavedDrive + WoodDeckSF + OpenPorchSF + EnclosedPorch + `3SsnPorch` + ScreenPorch + PoolArea + log(MiscVal) + MoSold + YrSold + SaleCondition, data = housesTrain)
summary(fit)

#control = trainControl(method = "LOOCV", number = 10)
control = trainControl(method = "cv")
customModel = train(SalePrice ~ ., data = housesTrain, method = "lm", trControl = control)
summary(customModel)

results = housesTest
results$SalePrice = exp(predict(customModel, newdata = housesTest))
results = results |>
  select(Id, SalePrice)

# Export
write_csv(results, "Submissions/Custom_Model.csv")
